---
title: "BCS analyses"
author: "Kelly Sansom"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
  word_document:
    toc: true
---



```{r setup, echo=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(viewer = NULL)
library(haven)
library(dplyr)
library(labelled)
library(gtsummary)
library(naniar)
library(ggplot2)
library(haven)
library(labelled)
library(magick)
library(forcats)
library(survey)
library(ggplot2)
library(dplyr)
library(broom)
library(kableExtra)
library(tidyr)
library(ggplot2)
library(cowplot)
library(gridExtra)
library(grid)  # for unit()

```

```{r, echo=FALSE, warning=FALSE, include=FALSE, message = FALSE}
library(dplyr)
library(labelled)
library(gtsummary)
library(survey)
# load data
df1 <- readRDS("bcs1970_clean_Feb26.rds")
df1_dictionary <- look_for(df1)
# select only required variables
df1 <-  df1 %>% select( bcsid, y26_sex = cmsex, age_left_ft_ed = 	
actagel2,sc,jdemand1, marstat2, jdemand2, jdemand3, bmi_26, bmi_30, bmi_42, bmi_46, exp_time_30, employment_42y, employment_46y, exp_time_42, exp_time_46, exp_cuml_42, exp_cuml_cat_42, exp_cuml_46, exp_cuml_cat_46, date_30y, date_42y, date_46y,malaise_26y, malaise_30y,  malaise_42y, malaise_46y , depression_30, depression_42, depression_46, censor_42, censor_46, censor, sw_30, sw_42,sw_46, genhealth_26, genhealth_30, genhealth_42, genhealth_46 )


df1 <- df1 %>%
  ungroup() 



```

# Study {.tabset}

## Participants

```{r, echo=FALSE, warning=FALSE}

#g <- readRDS("C:/Users/kelly/OneDrive - Flinders/Murthy/R workbooks/BCS 1970/consort.rds")
#plot(g)

image_read("bcs1970_consort_diagram.png")


```
<br><br><br>

*Missing outcome and covariates required for modelling at baseline*
```{r}

df_missing_baseline <- df1 %>%
  mutate(
    across(
      c(y26_sex, sc, genhealth_30),
      ~ forcats::as_factor(.x)
    )
  ) %>%
  mutate(genhealth_30 = forcats::fct_drop(genhealth_30))

# fix labels for presentation

df_missing_baseline <- df_missing_baseline %>%
  set_variable_labels(
    y26_sex = "Sex",
    age_left_ft_ed = "Age left full-time education, years",
    sc = "Job type",
    bmi_30 = "Body mass index (kg/m^2)",
    malaise_30y = "Psychological distress",
    depression_30 = "Help-seeking for depressive symptoms"
  )

table_missing_30fu  <- df_missing_baseline %>% select(y26_sex, age_left_ft_ed, sc,  bmi_30,  genhealth_30, malaise_30y,  depression_30  ) %>% tbl_summary()

gtsummary::theme_gtsummary_compact()

table_missing_30fu 
```

*Missing outcome and covariates required for modelling at 42 year follow-up*
```{r}
df_missing_42fu <- df1 %>% filter (censor_42 ==0) %>%
  mutate(
    across(
      c( genhealth_42),
      ~ forcats::as_factor(.x)
    )
  ) %>%
  mutate(genhealth_42 = forcats::fct_drop(genhealth_42))

# fix labels for presentation

df_missing_42fu <- df_missing_42fu %>%
  set_variable_labels(
    bmi_42 = "Body mass index (kg/m^2)",
    malaise_42y = "Psychological distress",
    depression_42 = "Help-seeking for depressive symptoms"
  )

table_missing_42fu   <- df_missing_42fu %>% select( bmi_42, genhealth_42, malaise_42y,  depression_42  ) %>% tbl_summary( )
gtsummary::theme_gtsummary_compact()
table_missing_42fu 

```


```{r, echo=FALSE, warning=FALSE}

#Imputating missing variables


# Function
set.seed(123)
random_sample_imputation <- function(df, missing_vars, gender_var) {
  # Separate complete and incomplete cases
  complete_cases <- df %>% filter(complete.cases(.))
  incomplete_cases <- df %>% filter(!complete.cases(.))
  
  
  # Impute missing values
  for (i in 1:nrow(incomplete_cases)) {
    # Get the gender of the current incomplete case
    current_gender <- incomplete_cases %>% slice(i) %>% pull(!!sym(gender_var))
    
    # Filter complete cases by the same gender
    gender_filtered_cases <- complete_cases %>% filter(!!sym(gender_var) == current_gender)
    
    # Randomly sample a complete case from the filtered cases
    sampled_case <- gender_filtered_cases %>% sample_n(1)
    
    # Impute missing values
    for (var in missing_vars) {
      if (is.na(incomplete_cases[i, var])) {
        incomplete_cases[i, var] <- sampled_case[[var]]
      }
    }
  }
  
  # Combine the imputed incomplete cases with the original complete cases
  imputed_df <- bind_rows(complete_cases, incomplete_cases)
  
  return(imputed_df)
}

# Imputing 26 with gender filter
df_26_missing <- df1 %>% select(bcsid, y26_sex, bmi_26, genhealth_26, malaise_26y)
missing_vars <- c("bmi_26", "genhealth_26", "malaise_26y")
df26_imputed <- random_sample_imputation(df_26_missing, missing_vars, "y26_sex")


# Imputing 30 with gender filter
df_30_missing <- df1 %>% select(bcsid, y26_sex, bmi_30, genhealth_30, age_left_ft_ed, sc, malaise_30y)
missing_vars <- c("bmi_30", "genhealth_30", "age_left_ft_ed", "sc", "malaise_30y")
df30_imputed <- random_sample_imputation(df_30_missing, missing_vars, "y26_sex")

# Imputing 42 with gender filter
df_42_missing <- df1 %>% select(bcsid, y26_sex, bmi_42, genhealth_42, malaise_42y, depression_42)
missing_vars <- c("bmi_42", "genhealth_42", "malaise_42y", "depression_42")
df42_imputed <- random_sample_imputation(df_42_missing, missing_vars, "y26_sex")



df1_imputed <- full_join( df30_imputed, df26_imputed)
df1_imputed <- full_join( df1_imputed, df42_imputed)



df1_imputed <- left_join(df1_imputed, df1, by = join_by("bcsid"))
# Only keep .x
df1_imputed <- df1_imputed %>%
  select(-ends_with(".y")) %>%
  rename_with(~ gsub("\\.x$", "", .), ends_with(".x"))

df1_complete <- df1_imputed %>% select(bcsid,y26_sex, age_left_ft_ed,sc,jdemand1, marstat2, jdemand2, jdemand3, bmi_26, bmi_30, bmi_42, bmi_46, exp_time_30, exp_time_42, exp_time_46, exp_cuml_42, exp_cuml_cat_42, exp_cuml_46, exp_cuml_cat_46, date_30y, date_42y, date_46y, malaise_26y, malaise_30y,  malaise_42y, malaise_46y , depression_30, depression_42, depression_46, censor_42, censor_46,censor, sw_30, sw_42,sw_46,genhealth_26, genhealth_30, genhealth_42, genhealth_46)

saveRDS(df1_complete, file = "df1_complete_46.rds")


```

Sample Baseline Table 1 at 30-year follow-up
```{r,}


df1_complete_clean <- df1_complete %>%
  mutate(
    across(
      c(y26_sex, sc, malaise_30y, genhealth_30, jdemand1, jdemand2, jdemand3, marstat2),
      ~ forcats::as_factor(.x)
    )
  ) %>%
  mutate(genhealth_30 = forcats::fct_drop(genhealth_30))

df1_complete_clean$sw_30 <- factor(df1_complete$sw_30)
# fix labels for presentation

df1_complete_clean <- df1_complete_clean %>%
  set_variable_labels(
    y26_sex = "Sex",
    age_left_ft_ed = "Age left full-time education, years",
    sc = "Job type",
    bmi_30 = "Body mass index (kg/m^2)",
    malaise_30y = "Psychological distress",
    depression_30 = "Help-seeking for depressive symptoms"
  )

table_1_baseline  <- df1_complete_clean %>% select(y26_sex, age_left_ft_ed,marstat2, sc,  bmi_30, genhealth_30,jdemand1, jdemand2, jdemand3, sw_30 ) %>% tbl_summary(by = "sw_30" )
gtsummary::theme_gtsummary_compact()
table_1_baseline

```

Individuals censored at 42-year and 46-year follow-up

```{r}

df1 %>%
  filter(censor_42 == 1) %>%
  count(employment_42y, name = "n") %>%
  arrange(desc(n))


```
Shift work patterns 42
```{r}

library(dplyr)

df1 %>%
  filter(censor_42 == 0) %>%  # keep only uncensored
  mutate(pattern = case_when(
    sw_30 == 1 & sw_42 == 0 ~ "Shift work at 30 only",
    sw_30 == 0 & sw_42 == 1 ~ "Shift work first time at 42",
    sw_30 == 1 & sw_42 == 1 ~ "Shift work at both",
    TRUE ~ "No shift work"
  )) %>%
  count(pattern)

```
Distribution of psychological distress and help-seeking for depressive symptoms among exposure groups

```{r}


# Shared aesthetics (colors, linetypes, shapes)
cols <- c(
  "No shift work"               = "#CBD5E1",  # soft gray-blue
  "Shift work at 30 only"       = "#9CC3E6",  # light blue
  "Shift work first time at 42" = "#3C7FB1",  # medium blue
  "Shift work at both"          = "#0A4C86"   # dark blue
)
lt <- c(
  "No shift work"               = "solid",
  "Shift work at 30 only"       = "dashed",
  "Shift work first time at 42" = "dotdash",
  "Shift work at both"          = "solid"
)
sh <- c(
  "No shift work"               = 16, # circle
  "Shift work at 30 only"       = 17, # triangle
  "Shift work first time at 42" = 15, # square
  "Shift work at both"          = 19  # larger circle
)

# -------- Function --------
make_outcome_plot_table <- function(df, var30, var42, outcome_title = "Outcome", abbreviate_legend = FALSE) {
  # Map the original column names to display labels
  label_map <- setNames(c("Age 30", "Age 42"), c(var30, var42))

  # 1) Summarise prevalence (uncensored only)
  df_plot <- df %>%
    filter(censor_42 == 0) %>%
    mutate(
      pattern = case_when(
        sw_30 == 1 & sw_42 == 0 ~ "Shift work at 30 only",
        sw_30 == 0 & sw_42 == 1 ~ "Shift work first time at 42",
        sw_30 == 1 & sw_42 == 1 ~ "Shift work at both",
        TRUE ~ "No shift work"
      )
    ) %>%
    pivot_longer(
      cols = all_of(c(var30, var42)),     # select then recode names to Age 30 / Age 42
      names_to = "timepoint",
      values_to = "value"
    ) %>%
    mutate(timepoint = recode(timepoint, !!!label_map)) %>%
    group_by(pattern, timepoint) %>%
    summarise(
      n     = sum(!is.na(value)),
      cases = sum(value == 1, na.rm = TRUE),
      prop  = cases / n,
      .groups = "drop"
    ) %>%
    mutate(
      pattern   = factor(pattern,
                         levels = c("No shift work",
                                    "Shift work at 30 only",
                                    "Shift work first time at 42",
                                    "Shift work at both")),
      timepoint = factor(timepoint, levels = c("Age 30", "Age 42"))
    )

  # Optional shorter legend labels to save space
  legend_labels <- if (abbreviate_legend) {
    c("No SW", "SW at 30 only", "SW at 42 first", "SW at both")
  } else {
    names(cols)
  }

  # 2) Narrow plot with bottom-right inset legend; no gridlines
  p <- ggplot(df_plot, aes(x = timepoint, y = prop,
                           group = pattern, color = pattern,
                           linetype = pattern, shape = pattern)) +
    geom_line(linewidth = 1.1) +              # slightly thinner for compactness
    geom_point(size = 2.6) +                  # smaller points for narrow figure
    scale_color_manual(values = cols, name = "Shift work (SW) exposure", labels = legend_labels) +
    scale_linetype_manual(values = lt, guide = "none") +
    scale_shape_manual(values = sh, guide = "none") +
    scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                       limits = c(0, NA),
                       expand = expansion(mult = c(0.02, 0.06))) +
    labs(
      x = "Age",
      y = "Prevalence (%)",
      title = outcome_title
    ) +
    # theme_classic removes major/minor gridlines by default
    theme_classic(base_size = 11) +
    theme(
      # Inset legend at bottom-right the plot
      legend.position = c(0.98, 0.02),
      legend.justification = c("right", "bottom"),
      legend.background = element_rect(fill = scales::alpha("white", 0.8), color = NA),
      legend.key.size = unit(6, "pt"),
      legend.text = element_text(size = 8),
      legend.title = element_text(size = 9, face = "bold"),
      # Compact margins
      plot.title = element_text(face = "bold", size = 11, margin = margin(b = 4)),
      axis.title.x = element_text(margin = margin(t = 2)),
      axis.title.y = element_text(margin = margin(r = 4)),
      plot.margin = margin(t = 2, r = 6, b = 2, l = 6)
    )

  
  df_table <- df_plot %>%
    rowwise() %>%
    mutate(
      ci_low  = if (n > 0) prop.test(cases, n)$conf.int[1] else NA_real_,
      ci_high = if (n > 0) prop.test(cases, n)$conf.int[2] else NA_real_,
      cell    = paste0(
        scales::percent(prop, accuracy = 0.1), " (",
        scales::percent(ci_low, accuracy = 0.1), "–",
        scales::percent(ci_high, accuracy = 0.1), "); n=", n
      )
    ) %>%
    ungroup() %>%
    select(`Exposure pattern` = pattern, timepoint, cell) %>%
    tidyr::pivot_wider(names_from = timepoint, values_from = cell) %>%
    arrange(`Exposure pattern`)

  table_theme <- ttheme_minimal(
    base_size = 9,
    core = list(fg_params = list(hjust = 0, x = 0.02),
                padding = unit(c(2, 2), "mm")),
    colhead = list(fg_params = list(fontface = 2, hjust = 0, x = 0.02),
                   padding = unit(c(2, 2), "mm"))
  )
  tbl_grob <- tableGrob(df_table, rows = NULL, theme = table_theme)

  # 4) Stack 
  combo <- cowplot::plot_grid(
    p, tbl_grob,
    ncol = 1,
    rel_heights = c(1.3, 0.7)  # less space for table, compact overall
  )

  return(combo)
}



# Malaise
fig_malaise <- make_outcome_plot_table(
  df = df1_complete,
  var30 = "malaise_30y",
  var42 = "malaise_42y",
  outcome_title = "Psychological distress",
  abbreviate_legend = TRUE   # set FALSE if you want full labels
)

# Depression
fig_depress <- make_outcome_plot_table(
  df = df1_complete,
  var30 = "depression_30",
  var42 = "depression_42",
  outcome_title = "Help-seeking for depressive symptoms",
  abbreviate_legend = TRUE
)

# Stack 
mh_figure <- cowplot::plot_grid(
  fig_malaise, fig_depress,
  ncol = 1, rel_heights = c(1, 1),
  labels = c("A", "B"), label_size = 12
)

# Save 
ggsave("mh_plot_with_table_compact.png", plot = mh_figure,
       width = 5, height = 9, dpi = 300)

print(mh_figure)

```




## Creating weights {.tabset}

### Weights for whole sample 

IPTW
```{r}
### IPTW (treatment weights) -----------------------------------------------

# Weight 1: baseline exposure at 30
wt1_num <- glm(sw_30 ~ 1, data = df1_complete, family = binomial)
wt1_fv_num <- fitted(wt1_num)
wt1_fv_num <- ifelse(df1_complete$sw_30 == 1, wt1_fv_num, 1 - wt1_fv_num)

wt1_denom <- glm(sw_30 ~ y26_sex + age_left_ft_ed + sc + bmi_26 + genhealth_26 + malaise_26y  ,
                 data = df1_complete, family = binomial)
wt1_fv_denom <- fitted(wt1_denom)
wt1_fv_denom <- ifelse(df1_complete$sw_30 == 1, wt1_fv_denom, 1 - wt1_fv_denom)

wt1 <- wt1_fv_num / wt1_fv_denom


# Weight 2: exposure at 42
wt2_num <- glm(sw_42 ~ sw_30, data = df1_complete, family = binomial)
wt2_fv_num <- fitted(wt2_num)
wt2_fv_num <- ifelse(df1_complete$sw_42 == 1, wt2_fv_num, 1 - wt2_fv_num)

wt2_denom <- glm(sw_42 ~ sw_30 + y26_sex + age_left_ft_ed + sc + bmi_26 +   
                           bmi_30 +  genhealth_26 +genhealth_30 +  malaise_26y +malaise_30y + depression_30 + malaise_42y + depression_42, # include mh and depression at 42 if outocme is outcome is 46.
                 data = df1_complete, family = binomial)
wt2_fv_denom <- fitted(wt2_denom)
wt2_fv_denom <- ifelse(df1_complete$sw_42 == 1, wt2_fv_denom, 1 - wt2_fv_denom)

wt2 <- wt2_fv_num / wt2_fv_denom


```

```{r}

### IPCW (censoring weights) -----------------------------------------------
# people not eligible at 42 are censored.
# Define uncensoring indicators (1 = observed, 0 = censored)
uncens_42 = 0
df1_complete$uncens_42 <- 1 - df1_complete$censor_42

# Censor weight at 42
cwt2_num <- glm(uncens_42 ~ sw_30, data = df1_complete, family = binomial)
cwt2_p_num <- fitted(cwt2_num)

cwt2_denom <- glm(uncens_42 ~ sw_30 +
                              y26_sex + age_left_ft_ed + sc +
                              bmi_26 + bmi_30 + genhealth_26 + genhealth_30 + malaise_30y + depression_30,
                  data = df1_complete, family = binomial)
cwt2_p_denom <- fitted(cwt2_denom)

cwt2_num_p   <- ifelse(df1_complete$uncens_42 == 1, cwt2_p_num,   1 - cwt2_p_num)
cwt2_denom_p <- ifelse(df1_complete$uncens_42 == 1, cwt2_p_denom, 1 - cwt2_p_denom)

cwt2 <- cwt2_num_p / cwt2_denom_p

# censor indicator variables
# 1 = observed, 0 = missing → censored at 46
df1_complete$uncens_malaise46    <- ifelse(is.na(df1_complete$malaise_46y), 0, 1)
df1_complete$uncens_depression46 <- ifelse(is.na(df1_complete$depression_46), 0, 1)

#################################################################################
# Censor weight for missing at malaise at 46
#################################################################################

# Numerator: depends only on treatment history
cwt3_mal_num <- glm(uncens_malaise46 ~ sw_30 + sw_42,
                    data = df1_complete, family = binomial)
cwt3_mal_p_num <- fitted(cwt3_mal_num)

# Denominator: full history predicting missingness
cwt3_mal_denom <- glm(uncens_malaise46 ~ sw_30 + sw_42 +
                        y26_sex + age_left_ft_ed + sc +
                        bmi_26 + bmi_30 +
                        genhealth_26 + genhealth_30 +
                        malaise_26y + malaise_30y + depression_30 +
                        malaise_42y + depression_42,
                      data = df1_complete, family = binomial)
cwt3_mal_p_denom <- fitted(cwt3_mal_denom)

# Convert to stabilized probabilities
cwt3_mal_num_p   <- ifelse(df1_complete$uncens_malaise46 == 1, 
                           cwt3_mal_p_num, 1 - cwt3_mal_p_num)
cwt3_mal_denom_p <- ifelse(df1_complete$uncens_malaise46 == 1, 
                           cwt3_mal_p_denom, 1 - cwt3_mal_p_denom)

# Final censoring weight for malaise outcome
cwt3_malaise <- cwt3_mal_num_p / cwt3_mal_denom_p

#################################################################################
# Censor weight for missing at medical help-seeking for depressive symp at 46
#################################################################################
# Numerator
cwt3_dep_num <- glm(uncens_depression46 ~ sw_30 + sw_42,
                    data = df1_complete, family = binomial)
cwt3_dep_p_num <- fitted(cwt3_dep_num)

# Denominator
cwt3_dep_denom <- glm(uncens_depression46 ~ sw_30 + sw_42 +
                        y26_sex + age_left_ft_ed + sc +
                        bmi_26 + bmi_30 +
                        genhealth_26 + genhealth_30 +
                        malaise_26y + malaise_30y + depression_30 +
                        malaise_42y + depression_42,
                      data = df1_complete, family = binomial)
cwt3_dep_p_denom <- fitted(cwt3_dep_denom)

# Stabilized probabilities
cwt3_dep_num_p   <- ifelse(df1_complete$uncens_depression46 == 1, 
                           cwt3_dep_p_num, 1 - cwt3_dep_p_num)
cwt3_dep_denom_p <- ifelse(df1_complete$uncens_depression46 == 1, 
                           cwt3_dep_p_denom, 1 - cwt3_dep_p_denom)

# Final censoring weight for depression outcome
cwt3_depression <- cwt3_dep_num_p / cwt3_dep_denom_p

################################################################################
# Combine with censor at wt at 42 
##############################################################################


ipcw_malaise    <- cwt2 * cwt3_malaise
ipcw_depression <- cwt2 * cwt3_depression

################################################################################
# Final weight for each outcome 
################################################################################
# IPTW already computed as wt1 * wt2
iptw <- wt1 * wt2

# Combine
final_wt_malaise    <- iptw * ipcw_malaise
final_wt_depression <- iptw * ipcw_depression

# Truncate each at 95th percentile
thr_mal <- quantile(final_wt_malaise, 0.95, na.rm = TRUE)
thr_dep <- quantile(final_wt_depression, 0.95, na.rm = TRUE)

final_wt_malaise_trunc    <- ifelse(final_wt_malaise > thr_mal, thr_mal, final_wt_malaise)
final_wt_depression_trunc <- ifelse(final_wt_depression > thr_dep, thr_dep, final_wt_depression)

df1_complete$wt_malaise    <- final_wt_malaise_trunc
df1_complete$wt_depression <- final_wt_depression_trunc

##############################################################################

```



```{r}
df1_uncensored <- df1_complete %>% filter(uncens_42 ==1 )

#df1_uncensored %>% select(-bcsid) %>% tbl_summary()

# exposure
df1_uncensored$exp_pattern_42 <- ifelse(
  df1_uncensored$sw_30 == 1 & df1_uncensored$sw_42 == 1 , 2,
  ifelse(
    df1_uncensored$sw_30 == 0 & df1_uncensored$sw_42 == 0 , 0,
    1
  )
)

df1_uncensored$exp_pattern_42 <- factor(df1_uncensored$exp_pattern_42)



 # 0 no exposure, 1 intermittent and 2 cont expo up to 42 and 46
```
Covariate Balance love plots

```{r, eval = FALSE}
##############################################
# Covariate balance with cobalt (Love plots) #
##############################################


suppressPackageStartupMessages({
  library(cobalt)
  library(dplyr)
  library(ggplot2)
  library(patchwork)  #  combining plots
})

# Put weights in the data & ensure variable types 

df1_complete <- df1_complete %>%
  mutate(
                                   # stabilized IPTW (truncated)
    sw_30 = as.integer(sw_30),
   sw_42 = as.integer(sw_42),
    y26_sex = factor(y26_sex),
    sc   = factor(sc)  )

# Covariates ----
covs_30 <- c("y26_sex", "age_left_ft_ed", "sc", "bmi_26", "malaise_26y", "genhealth_26")
covs_42 <- c("y26_sex", "age_left_ft_ed", "sc", "genhealth_30", "bmi_30", "bmi_42", "genhealth_42", "malaise_30y", "bmi_26", "malaise_26y", "genhealth_26", "depression_30")

# love plot function
love_plot_time <- function(data, treat_var, covs, weight_var, time_label) {
  # Build T ~ X1 + X2 + ...
  fml <- as.formula(paste(treat_var, "~", paste(covs, collapse = " + ")))

  # complete rows on the variables used at this timepoint (incl. weights)
  rows_keep <- stats::complete.cases(data[, c(treat_var, covs, weight_var)])
  dat <- data[rows_keep, , drop = FALSE]

  # Balance table (unweighted vs weighted)
  bt <- bal.tab(
    x         = fml,
    data      = dat,
    weights   = dat[[weight_var]],
    estimand  = "ATE",
    s.d.denom = "pooled",
    un        = TRUE,
    quick     = FALSE
  )

  # Love plot
  p <- love.plot(
    bt,
    abs        = TRUE,
    thresholds = c(m = 0.1),
    var.order  = "unadjusted",
    colors     = c(Un = "#7f8c8d", Adj = "#2c3e50"),
    shapes     = c(Un = 16, Adj = 17),
    size       = 4,
    title      = paste0("Covariate Balance (", time_label, ")"),
    subtitle   = "Standardised Mean Differences (Unweighted vs IPTW‑Weighted)"
  )

  list(bt = bt, plot = p, n = nrow(dat))
}

# each follow-up
res30 <- love_plot_time(df1_complete, "sw_30", covs_30, "wt1", "Age 30y")
res42 <- love_plot_time(df1_complete, "sw_30", covs_42, "wt2", "Age 42y")



library(patchwork)
library(ggplot2)

# Combine into one figure with larger text and proper layout
combined_plot <- (res30$plot + res42$plot ) +
  plot_layout(nrow = 2) +
  plot_annotation(
    title = "Covariate Balance Among Shift vs Non‑Shift Workers",
    subtitle = "Love plots by follow-up (30y, 42y)",
    theme = theme(
      plot.title = element_text(face = "bold", size = 20),
      plot.subtitle = element_text(size = 16)
    )
  ) &
  theme(
    axis.title = element_text(size = 16),
    axis.text = element_text(size = 18),
    legend.title = element_text(size = 18),
    legend.text = element_text(size = 18)
  )

# combined plot
print(combined_plot)

# saved
ggsave("balance_cov_2cols.png", combined_plot, width = 14, height = 20, dpi = 300)


```

### Weights for females
```{r}

idx_female <- df1_complete$y26_sex == 2
df_f <- df1_complete[idx_female, ]
### ---------------- FEMALE DATA ----------------
df_f <- df1_complete[df1_complete$y26_sex == 2, ]


### ============================================================
### 1. IPTW (treatment at 30 and 42) — same structure as whole sample
### ============================================================

## Weight 1: SW30
f_wt1_num <- glm(sw_30 ~ 1, data = df_f, family = binomial)
f_wt1_p_num <- fitted(f_wt1_num)

f_wt1_denom <- glm(sw_30 ~ age_left_ft_ed + sc +
                     bmi_26 + genhealth_26 + malaise_26y,
                   data = df_f, family = binomial)
f_wt1_p_denom <- fitted(f_wt1_denom)

f_wt1 <- ifelse(df_f$sw_30 == 1, f_wt1_p_num, 1 - f_wt1_p_num) /
         ifelse(df_f$sw_30 == 1, f_wt1_p_denom, 1 - f_wt1_p_denom)


## Weight 2: SW42
# IMPORTANT: censor_42 must already code missing sw_42 as 1 in df1_complete
f_wt2_num <- glm(sw_42 ~ sw_30, data = df_f, family = binomial)
f_wt2_p_num <- fitted(f_wt2_num)

f_wt2_denom <- glm(sw_42 ~ sw_30 +
                     bmi_30 + genhealth_30 + malaise_30y +
                     age_left_ft_ed + sc +
                     bmi_26 + genhealth_26 + malaise_26y + depression_30 +
                     malaise_42y + depression_42,
                   data = df_f, family = binomial)
f_wt2_p_denom <- fitted(f_wt2_denom)

f_wt2 <- ifelse(df_f$sw_42 == 1, f_wt2_p_num, 1 - f_wt2_p_num) /
         ifelse(df_f$sw_42 == 1, f_wt2_p_denom, 1 - f_wt2_p_denom)

iptw_female <- f_wt1 * f_wt2

thr_iptw_female <- quantile(iptw_female, 0.95, na.rm = TRUE)
iptw_female <- ifelse(iptw_female > thr_iptw_female,
                      thr_iptw_female, iptw_female)



### ============================================================
### 2. IPCW at 42 (same structure as whole sample)
### ============================================================

df_f$uncens_42 <- 1 - df_f$censor_42

f_cwt2_num <- glm(uncens_42 ~ sw_30, data = df_f, family = binomial)
f_cwt2_p_num <- fitted(f_cwt2_num)

f_cwt2_denom <- glm(uncens_42 ~ sw_30 +
                      age_left_ft_ed + sc +
                      bmi_30 + genhealth_30 + malaise_30y + depression_30 ,
                    data = df_f, family = binomial)
f_cwt2_p_denom <- fitted(f_cwt2_denom)

ipcw_female_42 <- ifelse(df_f$uncens_42 == 1, f_cwt2_p_num,   1 - f_cwt2_p_num) /
                  ifelse(df_f$uncens_42 == 1, f_cwt2_p_denom, 1 - f_cwt2_p_denom)



### ============================================================
### 3. IPCW at 46 for malaise_46y
### ============================================================

df_f$uncens_malaise46 <- ifelse(is.na(df_f$malaise_46y), 0, 1)

f_cwt3_mal_num <- glm(uncens_malaise46 ~ sw_30 + sw_42,
                      data = df_f, family = binomial)
f_cwt3_mal_p_num <- fitted(f_cwt3_mal_num)

f_cwt3_mal_denom <- glm(uncens_malaise46 ~ sw_30 + sw_42 +
                          age_left_ft_ed + sc +
                          bmi_26 + bmi_30 +
                          genhealth_26 + genhealth_30 +
                          malaise_26y + malaise_30y + depression_30 +
                          malaise_42y + depression_42,
                        data = df_f, family = binomial)
f_cwt3_mal_p_denom <- fitted(f_cwt3_mal_denom)

ipcw_female_malaise46 <- ifelse(df_f$uncens_malaise46 == 1, f_cwt3_mal_p_num,   1 - f_cwt3_mal_p_num) /
                         ifelse(df_f$uncens_malaise46 == 1, f_cwt3_mal_p_denom, 1 - f_cwt3_mal_p_denom)



### ============================================================
### 4. IPCW at 46 for depression_46
### ============================================================

df_f$uncens_dep46 <- ifelse(is.na(df_f$depression_46), 0, 1)

f_cwt3_dep_num <- glm(uncens_dep46 ~ sw_30 + sw_42,
                      data = df_f, family = binomial)
f_cwt3_dep_p_num <- fitted(f_cwt3_dep_num)

f_cwt3_dep_denom <- glm(uncens_dep46 ~ sw_30 + sw_42 +
                          age_left_ft_ed + sc +
                          bmi_26 + bmi_30 +
                          genhealth_26 + genhealth_30 +
                          malaise_26y + malaise_30y + depression_30 +
                          malaise_42y + depression_42,
                        data = df_f, family = binomial)
f_cwt3_dep_p_denom <- fitted(f_cwt3_dep_denom)

ipcw_female_dep46 <- ifelse(df_f$uncens_dep46 == 1, f_cwt3_dep_p_num,   1 - f_cwt3_dep_p_num) /
                     ifelse(df_f$uncens_dep46 == 1, f_cwt3_dep_p_denom, 1 - f_cwt3_dep_p_denom)



### ============================================================
### 5. FINAL FEMALE WEIGHTS (same structure as whole sample)
### ============================================================

## Malaise
final_malaise_female <- iptw_female * ipcw_female_42 * ipcw_female_malaise46
thr_mal_f <- quantile(final_malaise_female, 0.95, na.rm = TRUE)
final_malaise_female <- ifelse(final_malaise_female > thr_mal_f,
                               thr_mal_f, final_malaise_female)

## Depression
final_depression_female <- iptw_female * ipcw_female_42 * ipcw_female_dep46
thr_dep_f <- quantile(final_depression_female, 0.95, na.rm = TRUE)
final_depression_female <- ifelse(final_depression_female > thr_dep_f,
                                  thr_dep_f, final_depression_female)


### Attach to df_f
df_f$final_wt_malaise_female    <- final_malaise_female
df_f$final_wt_depression_female <- final_depression_female
df_f$iptw_female                <- iptw_female
df_f$ipcw_female_42             <- ipcw_female_42
df_f$ipcw_female_malaise46      <- ipcw_female_malaise46
df_f$ipcw_female_dep46          <- ipcw_female_dep46

```


```{r}
df_f_uncensored <- df_f %>% filter(uncens_42 == 1)

# exposure
df_f_uncensored$exp_pattern_42 <- ifelse(
  df_f_uncensored$sw_30 == 1 & df_f_uncensored$sw_42  == 1, 2,
  ifelse(
    df_f_uncensored$sw_30 == 0 & df_f_uncensored$sw_42 == 0 , 0,
    1
  )
)

df_f_uncensored$exp_pattern_42 <- factor(df_f_uncensored$exp_pattern_42)


 # 0 no exposure, 1 intermittent and 2 cont expo up to 42 
```

### Weights for males

```{r}

## -------------------- MALAise — MALES --------------------
## Subset
idx_male <- df1_complete$y26_sex == 1
df_m <- df1_complete[idx_male, ]


### ============================================================
### 1. IPTW (treatment at 30 and 42)
### ============================================================

## Weight 1: SW30
m_wt1_num <- glm(sw_30 ~ 1, data = df_m, family = binomial)
m_wt1_p_num <- fitted(m_wt1_num)

m_wt1_denom <- glm(sw_30 ~ age_left_ft_ed + sc +
                     bmi_26 + genhealth_26 + malaise_26y,
                   data = df_m, family = binomial)
m_wt1_p_denom <- fitted(m_wt1_denom)

m_wt1 <- ifelse(df_m$sw_30 == 1, m_wt1_p_num, 1 - m_wt1_p_num) /
         ifelse(df_m$sw_30 == 1, m_wt1_p_denom, 1 - m_wt1_p_denom)


## Weight 2: SW42
# IMPORTANT: censor_42 must already code missing sw_42 as 1 in df1_complete
m_wt2_num <- glm(sw_42 ~ sw_30, data = df_m, family = binomial)
m_wt2_p_num <- fitted(m_wt2_num)

m_wt2_denom <- glm(sw_42 ~ sw_30 +
                     bmi_30 + genhealth_30 + malaise_30y +
                     age_left_ft_ed + sc +
                     bmi_26 + genhealth_26 + malaise_26y + depression_30 +
                     malaise_42y + depression_42,
                   data = df_m, family = binomial)
m_wt2_p_denom <- fitted(m_wt2_denom)

m_wt2 <- ifelse(df_m$sw_42 == 1, m_wt2_p_num, 1 - m_wt2_p_num) /
         ifelse(df_m$sw_42 == 1, m_wt2_p_denom, 1 - m_wt2_p_denom)

iptw_male <- m_wt1 * m_wt2

thr_iptw_male <- quantile(iptw_male, 0.95, na.rm = TRUE)
iptw_male <- ifelse(iptw_male > thr_iptw_male,
                    thr_iptw_male, iptw_male)



### ============================================================
### 2. IPCW at 42
### ============================================================

df_m$uncens_42 <- 1 - df_m$censor_42

m_cwt2_num <- glm(uncens_42 ~ sw_30, data = df_m, family = binomial)
m_cwt2_p_num <- fitted(m_cwt2_num)

m_cwt2_denom <- glm(uncens_42 ~ sw_30 +
                      age_left_ft_ed + sc +
                      bmi_30 + genhealth_30 + malaise_30y + depression_30,
                    data = df_m, family = binomial)
m_cwt2_p_denom <- fitted(m_cwt2_denom)

ipcw_male_42 <- ifelse(df_m$uncens_42 == 1, m_cwt2_p_num,   1 - m_cwt2_p_num) /
                ifelse(df_m$uncens_42 == 1, m_cwt2_p_denom, 1 - m_cwt2_p_denom)



### ============================================================
### 3. IPCW at 46 for malaise_46y
### ============================================================

df_m$uncens_malaise46 <- ifelse(is.na(df_m$malaise_46y), 0, 1)

m_cwt3_mal_num <- glm(uncens_malaise46 ~ sw_30 + sw_42,
                      data = df_m, family = binomial)
m_cwt3_mal_p_num <- fitted(m_cwt3_mal_num)

m_cwt3_mal_denom <- glm(uncens_malaise46 ~ sw_30 + sw_42 +
                          age_left_ft_ed + sc +
                          bmi_26 + bmi_30 +
                          genhealth_26 + genhealth_30 +
                          malaise_26y + malaise_30y + depression_30 +
                          malaise_42y + depression_42,
                        data = df_m, family = binomial)
m_cwt3_mal_p_denom <- fitted(m_cwt3_mal_denom)

ipcw_male_malaise46 <- ifelse(df_m$uncens_malaise46 == 1, m_cwt3_mal_p_num,   1 - m_cwt3_mal_p_num) /
                       ifelse(df_m$uncens_malaise46 == 1, m_cwt3_mal_p_denom, 1 - m_cwt3_mal_p_denom)



### ============================================================
### 4. IPCW at 46 for depression_46
### ============================================================

df_m$uncens_dep46 <- ifelse(is.na(df_m$depression_46), 0, 1)

m_cwt3_dep_num <- glm(uncens_dep46 ~ sw_30 + sw_42,
                      data = df_m, family = binomial)
m_cwt3_dep_p_num <- fitted(m_cwt3_dep_num)

m_cwt3_dep_denom <- glm(uncens_dep46 ~ sw_30 + sw_42 +
                          age_left_ft_ed + sc +
                          bmi_26 + bmi_30 +
                          genhealth_26 + genhealth_30 +
                          malaise_26y + malaise_30y + depression_30 +
                          malaise_42y + depression_42,
                        data = df_m, family = binomial)
m_cwt3_dep_p_denom <- fitted(m_cwt3_dep_denom)

ipcw_male_dep46 <- ifelse(df_m$uncens_dep46 == 1, m_cwt3_dep_p_num,   1 - m_cwt3_dep_p_num) /
                   ifelse(df_m$uncens_dep46 == 1, m_cwt3_dep_p_denom, 1 - m_cwt3_dep_p_denom)



### ============================================================
### 5. FINAL MALE WEIGHTS
### ============================================================

## Malaise
final_malaise_male <- iptw_male * ipcw_male_42 * ipcw_male_malaise46
thr_mal_m <- quantile(final_malaise_male, 0.95, na.rm = TRUE)
final_malaise_male <- ifelse(final_malaise_male > thr_mal_m,
                             thr_mal_m, final_malaise_male)

## Depression
final_depression_male <- iptw_male * ipcw_male_42 * ipcw_male_dep46
thr_dep_m <- quantile(final_depression_male, 0.95, na.rm = TRUE)
final_depression_male <- ifelse(final_depression_male > thr_dep_m,
                                thr_dep_m, final_depression_male)


### Attach to df_m
df_m$final_wt_malaise_male    <- final_malaise_male
df_m$final_wt_depression_male <- final_depression_male
df_m$iptw_male                <- iptw_male
df_m$ipcw_male_42             <- ipcw_male_42
df_m$ipcw_male_malaise46      <- ipcw_male_malaise46
df_m$ipcw_male_dep46          <- ipcw_male_dep46

```


```{r}
df_m_uncensored <- df_m %>% filter(uncens_42 == 1)

# exposure
df_m_uncensored$exp_pattern_42 <- ifelse(
  df_m_uncensored$sw_30 == 1 & df_m_uncensored$sw_42  == 1, 2,
  ifelse(
    df_m_uncensored$sw_30 == 0 & df_m_uncensored$sw_42 == 0 , 0,
    1
  )
)

df_m_uncensored$exp_pattern_42 <- factor(df_m_uncensored$exp_pattern_42)


 # 0 no exposure, 1 intermittent and 2 cont expo up to 42 
```


## Main outcome model {.tabset}


### main analysis
```{r}
### Whole population malaise
library(marginaleffects)
library(survey)
# Create a safe temp folder
#tempdir_path <- "C:/Users/kelly/Documents/BCS70_analysis/temp"
#dir.create(tempdir_path, showWarnings = FALSE, recursive = TRUE)

# Tell R to use this folder for temp files
#Sys.setenv(TMPDIR = tempdir_path)

# take 2

design_mal <- svydesign(ids = ~1, weights = ~ df1_uncensored$wt_malaise,
                        data = df1_uncensored)

# weighted counts


weighted_counts <- svytable(~exp_pattern_42, design_mal)
weighted_events <- svytable(~exp_pattern_42 + malaise_46y, design)[, "1"]

msm.mod1 <- svyglm(malaise_46y ~ exp_pattern_42 +  factor(y26_sex) + age_left_ft_ed +sc + bmi_26 + genhealth_26 + malaise_26y   , 
                  design = design_mal,
                  family = quasibinomial())


# ---------------------------------------------------------
# Robust (sandwich) variance–covariance matrix
# ---------------------------------------------------------
vcov_robust <- sandwich::vcovHC(msm.mod1, type = "HC0")

# ---------------------------------------------------------
# Absolute risks
# ---------------------------------------------------------
avg_predictions(
  msm.mod1,
  newdata = df1_uncensored,
  type = "response",
  wts = df1_uncensored$wt_malaise,
  by = "exp_pattern_42",
  vcov = vcov_robust
)

# ---------------------------------------------------------
# Risk differences
# ---------------------------------------------------------
risk_diff_main <- avg_comparisons(
  msm.mod1,
  variables = "exp_pattern_42",
  comparison = "difference",
  type = "response",
  newdata = df1_uncensored,
  wts = df1_uncensored$wt_malaise,
  vcov = vcov_robust
)

# ---------------------------------------------------------
# Relative risks
# ---------------------------------------------------------
risk_ratio_main <- avg_comparisons(
  msm.mod1,
  variables = "exp_pattern_42",
  comparison = "ratio",
  type = "response",
  newdata = df1_uncensored,
  wts = df1_uncensored$wt_malaise,
  vcov = vcov_robust
)


risk_diff_main

risk_ratio_main


```

```{r,}

# main population depressive symptoms
# Create a safe temp folder
#tempdir_path <- "C:/Users/kelly/Documents/BCS70_analysis/temp"
#dir.create(tempdir_path, showWarnings = FALSE, recursive = TRUE)




# ---------------------------------------------------------
# 1. Survey design for help-seeking outcome
# ---------------------------------------------------------
design_dep <- svydesign(
  ids = ~1,
  weights = ~ wt_depression, 
  data = df1_uncensored
)

# ---------------------------------------------------------
# 2. Fit the MSM for help-seeking
# ---------------------------------------------------------
mod_depression1 <- svyglm(
  depression_46 ~ exp_pattern_42 + factor(y26_sex) + age_left_ft_ed +sc + bmi_26 + genhealth_26 + malaise_26y,
  design = design_dep,
  family = quasibinomial()
)

# ---------------------------------------------------------
# 3. Robust (sandwich) variance–covariance matrix
# ---------------------------------------------------------
vcov_dep_robust <- sandwich::vcovHC(mod_depression1, type = "HC0")

# ---------------------------------------------------------
# 4. Absolute risks (predicted probabilities)
# ---------------------------------------------------------
abs_risks_dep <- avg_predictions(
  mod_depression1,
  newdata = df1_uncensored,
  type = "response",
  wts = df1_uncensored$wt_depression,
  by = "exp_pattern_42",
  vcov = vcov_dep_robust
)

# ---------------------------------------------------------
# 5. Risk differences
# ---------------------------------------------------------
risk_diff_dep_main <- avg_comparisons(
  mod_depression1,
  variables = "exp_pattern_42",
  comparison = "difference",
  type = "response",
  newdata = df1_uncensored,
  wts = df1_uncensored$wt_depression,
  vcov = vcov_dep_robust
)

# ---------------------------------------------------------
# 6. Relative risks
# ---------------------------------------------------------
risk_ratio_dep_main <- avg_comparisons(
  mod_depression1,
  variables = "exp_pattern_42",
  comparison = "ratio",
  type = "response",
  newdata = df1_uncensored,
  wts = df1_uncensored$wt_depression,
  vcov = vcov_dep_robust
)

# ---------------------------------------------------------
# 7. Print results
# ---------------------------------------------------------
abs_risks_dep
risk_diff_dep_main
risk_ratio_dep_main



```

### Females
malaise
```{r}
### Whole population malaise
library(marginaleffects)
# Create a safe temp folder
#tempdir_path <- "C:/Users/kelly/Documents/BCS70_analysis/temp"
#dir.create(tempdir_path, showWarnings = FALSE, recursive = TRUE)

# Tell R to use this folder for temp files
#Sys.setenv(TMPDIR = tempdir_path)
design <- svydesign(ids = ~1, weights = ~ final_wt_malaise_female, data =df_f_uncensored)

msm.mod1 <- svyglm(malaise_46y  ~ exp_pattern_42  + age_left_ft_ed +sc + bmi_26 + genhealth_26 + malaise_26y  , 
                  design = design, 
                  family = quasibinomial())

tbl <- tbl_regression(msm.mod1, exponentiate = TRUE) %>% add_glance_source_note() %>%   add_n(location = "level") %>% add_nevent(location = "level")



# ---------------------------------------------------------
# Robust (sandwich) variance–covariance matrix
# ---------------------------------------------------------
vcov_robust <- sandwich::vcovHC(msm.mod1, type = "HC0")

# ---------------------------------------------------------
# Absolute risks
# ---------------------------------------------------------
avg_predictions(
  msm.mod1,
  newdata =df_f_uncensored,
  type = "response",
  wts =df_f_uncensored$final_wt_malaise_female,
  by = "exp_pattern_42",
  vcov = vcov_robust
)

# ---------------------------------------------------------
# Risk differences
# ---------------------------------------------------------
risk_diff_malaise_female <- avg_comparisons(
  msm.mod1,
  variables = "exp_pattern_42",
  comparison = "difference",
  type = "response",
  newdata =df_f_uncensored,
  wts =df_f_uncensored$final_wt_malaise_female,
  vcov = vcov_robust
)

# ---------------------------------------------------------
# Relative risks
# ---------------------------------------------------------
risk_ratio_malaise_female <- avg_comparisons(
  msm.mod1,
  variables = "exp_pattern_42",
  comparison = "ratio",
  type = "response",
  newdata =df_f_uncensored,
  wts =df_f_uncensored$final_wt_malaise_female,
  vcov = vcov_robust
)
risk_diff_malaise_female
risk_ratio_malaise_female



```

depressive symptoms
```{r,}

# Create a safe temp folder
#tempdir_path <- "C:/Users/kelly/Documents/BCS70_analysis/temp"
#dir.create(tempdir_path, showWarnings = FALSE, recursive = TRUE)
df_f_uncensored$depression_42 <- as.factor(df_f_uncensored$depression_42)
# ---------------------------------------------------------
# 1. Survey design for help-seeking outcome
# ---------------------------------------------------------
design_dep <- svydesign(
  ids = ~1,
  weights = ~ final_wt_depression_female, 
  data = df_f_uncensored
)

# ---------------------------------------------------------
# 2. Fit the MSM for help-seeking
# ---------------------------------------------------------
mod_depression1 <- svyglm(
  depression_46 ~ exp_pattern_42 + age_left_ft_ed +sc + bmi_26 + genhealth_26 + malaise_26y,
  design = design_dep,
  family = quasibinomial()
)

# ---------------------------------------------------------
# 3. Robust (sandwich) variance–covariance matrix
# ---------------------------------------------------------
vcov_dep_robust <- sandwich::vcovHC(mod_depression1, type = "HC0")

# ---------------------------------------------------------
# 4. Absolute risks (predicted probabilities)
# ---------------------------------------------------------
abs_risks_dep <- avg_predictions(
  mod_depression1,
  newdata = df_f_uncensored,
  type = "response",
  wts = df_f_uncensored$final_wt_depression_female,
  by = "exp_pattern_42",
  vcov = vcov_dep_robust
)

# ---------------------------------------------------------
# 5. Risk differences
# ---------------------------------------------------------
risk_diff_dep_female <- avg_comparisons(
  mod_depression1,
  variables = "exp_pattern_42",
  comparison = "difference",
  type = "response",
  newdata = df_f_uncensored,
  wts = df_f_uncensored$final_wt_depression_female,
  vcov = vcov_dep_robust
)

# ---------------------------------------------------------
# 6. Relative risks
# ---------------------------------------------------------
risk_ratio_dep_female <- avg_comparisons(
  mod_depression1,
  variables = "exp_pattern_42",
  comparison = "ratio",
  type = "response",
  newdata = df_f_uncensored,
  wts = df_f_uncensored$final_wt_depression_female,
  vcov = vcov_dep_robust
)

# ---------------------------------------------------------
# 7. Print results
# ---------------------------------------------------------
risk_diff_dep_female
risk_ratio_dep_female



```

### Males

```{r}
### Whole population malaise
library(marginaleffects)
# Create a safe temp folder
#tempdir_path <- "C:/Users/kelly/Documents/BCS70_analysis/temp"
#dir.create(tempdir_path, showWarnings = FALSE, recursive = TRUE)

# Tell R to use this folder for temp files
#Sys.setenv(TMPDIR = tempdir_path)

design <- svydesign(ids = ~1, weights = ~ final_wt_malaise_male, data =df_m_uncensored)

msm.mod1 <- svyglm(malaise_46y  ~ exp_pattern_42  + age_left_ft_ed +sc + bmi_26 + genhealth_26 + malaise_26y  , 
                  design = design, 
                  family = quasibinomial())

tbl <- tbl_regression(msm.mod1, exponentiate = TRUE) %>% add_glance_source_note() %>%   add_n(location = "level") %>% add_nevent(location = "level")



# ---------------------------------------------------------
# Robust (sandwich) variance–covariance matrix
# ---------------------------------------------------------
vcov_robust <- sandwich::vcovHC(msm.mod1, type = "HC0")

# ---------------------------------------------------------
# Absolute risks
# ---------------------------------------------------------
avg_predictions(
  msm.mod1,
  newdata =df_m_uncensored,
  type = "response",
  wts =df_m_uncensored$final_wt_malaise_male,
  by = "exp_pattern_42",
  vcov = vcov_robust
)

# ---------------------------------------------------------
# Risk differences
# ---------------------------------------------------------
risk_diff_malaise_male <- avg_comparisons(
  msm.mod1,
  variables = "exp_pattern_42",
  comparison = "difference",
  type = "response",
  newdata =df_m_uncensored,
  wts =df_m_uncensored$final_wt_malaise_male,
  vcov = vcov_robust
)

# ---------------------------------------------------------
# Relative risks
# ---------------------------------------------------------
risk_ratio_malaise_male <- avg_comparisons(
  msm.mod1,
  variables = "exp_pattern_42",
  comparison = "ratio",
  type = "response",
  newdata =df_m_uncensored,
  wts =df_m_uncensored$final_wt_malaise_male,
  vcov = vcov_robust
)
risk_diff_malaise_male
risk_ratio_malaise_male



```

depressive symptoms
```{r,}

# Create a safe temp folder
tempdir_path <- "C:/Users/kelly/Documents/BCS70_analysis/temp"
#dir.create(tempdir_path, showWarnings = FALSE, recursive = TRUE)
df_m_uncensored$depression_42 <- as.factor(df_m_uncensored$depression_42)
# ---------------------------------------------------------
# 1. Survey design for help-seeking outcome
# ---------------------------------------------------------
design_dep <- svydesign(
  ids = ~1,
  weights = ~ final_wt_depression_male, 
  data = df_m_uncensored
)

# ---------------------------------------------------------
# 2. Fit the MSM for help-seeking
# ---------------------------------------------------------
mod_depression1 <- svyglm(
  depression_46 ~ exp_pattern_42 + age_left_ft_ed +sc + bmi_26 + genhealth_26 +malaise_26y,
  design = design_dep,
  family = quasibinomial()
)

# ---------------------------------------------------------
# 3. Robust (sandwich) variance–covariance matrix
# ---------------------------------------------------------
vcov_dep_robust <- sandwich::vcovHC(mod_depression1, type = "HC0")

# ---------------------------------------------------------
# 4. Absolute risks (predicted probabilities)
# ---------------------------------------------------------
abs_risks_dep <- avg_predictions(
  mod_depression1,
  newdata = df_m_uncensored,
  type = "response",
  wts = df_m_uncensored$final_wt_depression_male,
  by = "exp_pattern_42",
  vcov = vcov_dep_robust
)

# ---------------------------------------------------------
# 5. Risk differences
# ---------------------------------------------------------
risk_diff_dep_male <- avg_comparisons(
  mod_depression1,
  variables = "exp_pattern_42",
  comparison = "difference",
  type = "response",
  newdata = df_m_uncensored,
  wts = df_m_uncensored$final_wt_depression_male,
  vcov = vcov_dep_robust
)

# ---------------------------------------------------------
# 6. Relative risks
# ---------------------------------------------------------
risk_ratio_dep_male <- avg_comparisons(
  mod_depression1,
  variables = "exp_pattern_42",
  comparison = "ratio",
  type = "response",
  newdata = df_m_uncensored,
  wts = df_m_uncensored$final_wt_depression_male,
  vcov = vcov_dep_robust
)

# ---------------------------------------------------------
# 7. Print results
# ---------------------------------------------------------
risk_ratio_dep_male
risk_diff_dep_male



```




### Table final result summary 

```{r, warning=FALSE, error=FALSE, eval = FALSE, echo=FALSE}
library(dplyr)
library(flextable)

library(dplyr)
library(flextable)
library(marginaleffects)
library(survey)
# Create a safe temp folder
#tempdir_path <- "C:/Users/kelly/Documents/BCS70_analysis/temp"
#dir.create(tempdir_path, showWarnings = FALSE, recursive = TRUE)

# =========================================================
# 1. COUNTS BY SUBGROUP
# =========================================================

# Main sample
counts_main <- df1_uncensored %>%
  group_by(exp_pattern_42) %>%
  summarise(
    N = n(),
    distress_events = sum(malaise_46y == 1, na.rm = TRUE),
    help_events = sum(depression_46 == 1, na.rm = TRUE),
    .groups = "drop"
  ) %>% mutate(exp_pattern_42 = as.character(exp_pattern_42))

# Female sample
counts_female <- df_f_uncensored %>%
  group_by(exp_pattern_42) %>%
  summarise(
    N = n(),
    distress_events = sum(malaise_46y == 1, na.rm = TRUE),
    help_events = sum(depression_46 == 1, na.rm = TRUE),
    .groups = "drop"
  ) %>% mutate(exp_pattern_42 = as.character(exp_pattern_42))

# Male sample
counts_male <- df_m_uncensored %>%
  group_by(exp_pattern_42) %>%
  summarise(
    N = n(),
    distress_events = sum(malaise_46y == 1, na.rm = TRUE),
    help_events = sum(depression_46 == 1, na.rm = TRUE),
    .groups = "drop"
  ) %>% mutate(exp_pattern_42 = as.character(exp_pattern_42))


# =========================================================
# 2. HELPER FUNCTION TO FORMAT RD/RR
# =========================================================

format_effects <- function(effects_df, digits = 2) {
  effects_df %>%
    mutate(
      Exposure = case_when(
        grepl("^0$", contrast) ~ "0",
        grepl("^1$", contrast) ~ "1",
        grepl("^2$", contrast) ~ "2",
        grepl("^\\s*1\\s*-\\s*0\\s*$", contrast) ~ "1",  # RD style
        grepl("^\\s*2\\s*-\\s*0\\s*$", contrast) ~ "2",  # RD style
        grepl("mean\\(1\\)/mean\\(0\\)", contrast) ~ "1", # RR style
        grepl("mean\\(2\\)/mean\\(0\\)", contrast) ~ "2", # RR style
        TRUE ~ as.character(contrast)
      ),
      Exposure = as.character(Exposure),
      estimate = sprintf(paste0("%.", digits, "f"), estimate),
      conf.low = sprintf(paste0("%.", digits, "f"), conf.low),
      conf.high = sprintf(paste0("%.", digits, "f"), conf.high),
      stat = paste0(estimate, " (", conf.low, "–", conf.high, ")")
    ) %>%
    select(Exposure, stat)
}


# =========================================================
# 3. MAIN MSM RD + RR (already computed earlier)
# =========================================================

rd_distress_main <- format_effects(risk_diff_main, digits = 3) %>% rename(RD = stat)
rr_distress_main <- format_effects(risk_ratio_main, digits = 2) %>% rename(RR = stat)

rd_help_main <- format_effects(risk_diff_dep_main, digits = 3) %>% rename(RD = stat)
rr_help_main <- format_effects(risk_ratio_dep_main, digits = 2) %>% rename(RR = stat)


# =========================================================
# 4. FEMALE MSM MODELS (log link)
# =========================================================

# distress
rd_distress_female <- format_effects(risk_diff_malaise_female, digits = 3) %>% rename(RD = stat)
rr_distress_female <- format_effects(risk_ratio_malaise_female, digits = 2) %>% rename(RR = stat)

# Help-seeking

rd_help_female <- format_effects(risk_diff_dep_female, digits = 3) %>% rename(RD = stat)
rr_help_female <- format_effects(risk_ratio_dep_female, digits = 2) %>% rename(RR = stat)


# =========================================================
# 5. MALE MSM MODELS (log link)
# =========================================================
# distress
rd_distress_male <- format_effects(risk_diff_malaise_male, digits = 3) %>% rename(RD = stat)
rr_distress_male <- format_effects(risk_ratio_malaise_male, digits = 2) %>% rename(RR = stat)

# Help-seeking

rd_help_male <- format_effects(risk_diff_dep_male, digits = 3) %>% rename(RD = stat)
rr_help_male <- format_effects(risk_ratio_dep_male, digits = 2) %>% rename(RR = stat)


# =========================================================
# 6. BLOCK BUILDER
# =========================================================

# =========================================================
# 4) BLOCK BUILDER (one row per exposure: 0,1,2)
# =========================================================
make_block <- function(rd_tbl, rr_tbl, counts_df, label, event_var) {
  # Normalize keys then merge RD & RR by Exposure
  rd_tbl <- rd_tbl %>% mutate(Exposure = normalize_exposure_label(Exposure))
  rr_tbl <- rr_tbl %>% mutate(Exposure = normalize_exposure_label(Exposure))
  merged <- full_join(rd_tbl, rr_tbl, by = "Exposure")

  # Add reference row if missing
  if (!"0" %in% merged$Exposure) {
    merged <- bind_rows(tibble(Exposure = "0", RD = "—", RR = "—"), merged)
  }

  # Attach counts by normalized key
  merged <- merged %>%
    left_join(
      counts_df %>%
        mutate(exp_pattern_42 = as.character(exp_pattern_42)) %>%
        rename(Exposure = exp_pattern_42),
      by = "Exposure"
    ) %>%
    mutate(
      Subgroup = label,
      Events = .data[[event_var]],
      `Exposure pattern` = dplyr::case_when(
        Exposure == "0" ~ "0 (reference)",
        Exposure == "1" ~ "1 / 0",
        Exposure == "2" ~ "2 / 0",
        TRUE ~ Exposure
      ),
      Exposure_ord = dplyr::case_when(
        Exposure == "0" ~ 0L,
        Exposure == "1" ~ 1L,
        Exposure == "2" ~ 2L,
        TRUE ~ 99L
      )
    ) %>%
    arrange(Exposure_ord) %>%
    select(Subgroup, `Exposure pattern`, N, Events, RD, RR)

  merged
}


# =========================================================
# 5) BUILD BOTH TABLES (distress + help)
# =========================================================
distress_tbl <- bind_rows(
  make_block(rd_distress_main,   rr_distress_main,   counts_main,   "Main MSM",   "distress_events"),
  make_block(rd_distress_female, rr_distress_female, counts_female, "Female MSM", "distress_events"),
  make_block(rd_distress_male,   rr_distress_male,   counts_male,   "Male MSM",   "distress_events")
)

help_tbl <- bind_rows(
  make_block(rd_help_main,   rr_help_main,   counts_main,   "Main MSM",   "help_events"),
  make_block(rd_help_female, rr_help_female, counts_female, "Female MSM", "help_events"),
  make_block(rd_help_male,   rr_help_male,   counts_male,   "Male MSM",   "help_events")
)


# =========================================================
# 6) FLEXS — now both ft_distress and ft_help exist
# =========================================================
ft_distress <- flextable(distress_tbl) %>%
  set_header_labels(
    Subgroup = "Subgroup",
    `Exposure pattern` = "Exposure pattern",
    N = "N",
    Events = "Events",
    RD = "RD (95% CI)",
    RR = "RR (95% CI)"
  ) %>%
  bold(part = "header") %>%
  autofit()

ft_help <- flextable(help_tbl) %>%
  set_header_labels(
    Subgroup = "Subgroup",
    `Exposure pattern` = "Exposure pattern",
    N = "N",
    Events = "Events",
    RD = "RD (95% CI)",
    RR = "RR (95% CI)"
  ) %>%
  bold(part = "header") %>%
  autofit()

# Display
ft_distress
ft_help

save_as_docx(
  "Psychological distress" = ft_distress,
  "Help-seeking" = ft_help,
  path = "msm_results_tables_46.docx"
)



```


